---
title: "Snowshoe Report"
author: "Sachiko Lamen"
date: "11/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(broom)
library(equatiomatic)
library(lubridate)
library(kableExtra)
library(effsize)
```
# Add introduction A 4 - 5 sentence "mini-introduction" to what this study is exploring, including basic background information about the subject, site and study (with citations as needed), and a brief overview of what is included in the exploratory report. You should probably do this last so you know what’s contained in the report.

# Include data analysis A 3 - 4 sentence section in which you briefly describe the data, including citation & credit as needed, and provide an overview of the main analysis, statistical methods and tools used.

This report visualizes and wrangles data for juvenile Snowshoe Hares recorded from 1998 - 2012. Data was gathered from Bonanza Creek LTER as part of the US Long Term Ecological Research Network.

**Citation:** Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative. https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14

```{r}
bonanza <- read_csv(here("data", "bonanza_hares.csv"))
```

## Juvenile Hare Investigation
Create new data set that counts total number of juvenile hare trappings per year, then plot using `geom_line()` to visualize fluctuations in juvenile hare trappings per year from 1998 - 2012. 
```{r}
bonanza_year <- bonanza %>% # create new data set 
  mutate(date_new = mdy(date)) %>% 
  mutate(year = year(date_new)) %>% # with date converted to class "Date"
  filter(age == "j") %>% # only pull results where age = j (juvenile)
  select(year) %>% # select column `year`
  count(year) # counting total number of trappings per year

```

```{r}
ggplot(data = bonanza_year, aes(x = year)) +
  geom_line(aes(x = year, y = n)) +
  labs (x = "Year",
        y = "Number of Trappings",
        title = "Number of juvenile hares trapped per year (1998 - 2012)")
```
### Describe major takeaways from the graph and total snowshoe hare trappings at the sight (include information from summary table below (hide summary table)

```{r}
stats_bonanza <- bonanza_year %>%
  group_by(year) %>%
  summarize(total_trappings = sum(n))

stats_bonanza %>%
  summarize(mean = mean(total_trappings),
            median = median(total_trappings),
            std_deviation = sd(total_trappings),
            maximum = max(total_trappings),
            minimum = min(total_trappings))
```
## Visualize juvenile hare weights

```{r}
juveniles <- bonanza %>%
  filter(age %in% "j") %>%
           select(weight, sex, age, grid) %>%
  mutate(sex = case_when(sex == "m" ~ "Male",
                          sex == "f" ~ "Female"),
         grid = case_when(grid == "bonrip" ~ "Bonanza Riparian",
                          grid == "bonmat" ~ "Bonanza Mature",
                          grid == "bonbs" ~ "Lowland Black Spruce"))


ggplot(data = juveniles, aes(x = sex, y = weight)) +
  geom_jitter(aes(color = sex), width = 0.5, size = 0.6) +
  facet_wrap(~ grid) +
  labs(title = "Juvenile Hare Weight Distribution Across Sites",
       y = "Weight (g)") +
  theme(axis.title.x = element_blank())
```


## Juvenile weight comparison (male & female snowshoe hares)
Finalized table containing th emean, std dev, and sample size w/ table caption placed ABOVE the table 
```{r}

juveniles_stats <- juveniles %>%
  group_by(sex) %>%
  summarize(trappings = n(),
            mean = mean(trappings),
            median = median(trappings),
            std_deviation = sd(trappings),
            maximum = max(trappings),
            minimum = min(trappings))

juveniles_stats %>%
  kable(col.names = c("Sex",
                      "Sample Size",
                      "Mean Size (g)",
                      "Median Size (g)",
                      "Standard Deviation",
                      "Maximum Size (g)",
                      "Minimum Size (g)"),
        caption = "**Table 1. Male v. Female Statistics for Purple Urchins (Naples Reef, 2010).** Data was collected from Naples Reef (NAPL) for two treatments: CONTROL (n = 112) treatments left giant kelp intact, ANNUAL (n = 135) treatments removed giant kelp each winter. Urchin size was measured using diameter (cm)") %>%
  kable_styling(full_width = FALSE)

```

2-3 sentence comprehensive statement of the means comparison including: actual difference, effect size (cohen's d), and outcome of an unpaired two sample t test

```{r, include = FALSE}
ggplot(data = juveniles, aes(x = weight)) +
  geom_histogram(bins = 15) +
  facet_wrap(~ sex)

ggplot(data = juveniles, aes(sample = weight)) +
  geom_qq() +
  facet_wrap(~sex)
```


```{r}
juvenile_females <- juveniles %>% 
  filter(sex == "Female") %>% 
  pull(weight)

juvenile_males <- juveniles %>%
  filter(sex == "Male") %>%
  pull(weight)

juvenile_t <- t.test(juvenile_females, juvenile_males, na.rm = TRUE)

juvenile_t

juvenile_cohen<- cohen.d(juvenile_females, juvenile_males, na.rm = TRUE)

juvenile_cohen
```

## Relationship between juvenile weight and hind foot length

```{r}
foot_weight <- bonanza %>%
  filter(age == "j") %>%
  select(age, hindft, weight)

ggplot(data = foot_weight, aes(x = weight, y = hindft)) +
  geom_point() +
  labs(title = "Hind Foot Length and Weight Comparison",
       x = "Weight (g)",
       y = "Hind Foot (mm)")
```
## FIGURE CAPTION

## Simple Linear Regression
```{r}
foot_weight_lm <- lm(hindft~ weight, data = foot_weight)

summary(foot_weight_lm)
```








Sara Guitinan Bureau of Ocean Energy Management (BOEM): Fish species distributions are changing -- BOEM -- Morro Bay offshore wind planning area -- fishing industry needs to adapt to change in species distribution -- docks, onshore processing places-- what are the needs of the fishing industry -- support for filling the needs would be mitigation of the negative impacts of an offshore wind project
What would a osw project look like if it was centered around EJ
Something on green hydrogen? -- offshore winds could potentially be used to produce hydrogen -- electrolysis. Unknowns?

Near Diablo Canyon Nuclear site that will be decomissioned soon. Conflict of negotiations between DOD and someone else. Also important to inigenous people in the area. 

Involved in 4 GP
2019 - lifecycle assessment of greenhouse gas emmisions from floating offshore winds in CA
 - report is helping to inform offshore renewable energy
Marine mammal monitoring
Citizen science for LT rocky intertidal monitoring
Ocean Acidification -- monitoring gaps and hotspots

West Coast Ocean Alliance