---
title: "Snowshoe Report"
author: "Sachiko Lamen"
date: "11/4/2021"
output: 
  html_document: 
    theme: cerulean
    highlight: pygments
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(equatiomatic)
library(lubridate)
library(kableExtra)
library(effsize)
library(plotly)
library(broom)
library(ggbeeswarm)
```
**Introduction:** This report provides an exploratory overview of differences in juvenile snowshoe hares trapped in Bonanza Creek LTER (1999 - 2012). Hare weight is compared amongst sexes within and across three trapping sites (Bonanza Black Spruce, Bonanza Mature, and Bonanza Riparian). A simple linear regression model is used to understand the relationship between hind foot length and weight, not taking in to account differences in sex. 

**Data and Analysis** Juvenile snowshoe hare weight measurements were collected and made available by Dr. Kiellan Knut and colleagues at the Bonanza Creek Long Term Ecological Research (LTER) site in Alaska. The data contains observations for 3197 observations, collected from 1999 - 2012. Following exploratory data visualization, weight between male and female hares are by two-sample t-tests using a significance level (\(alpha\)) of 0.05 throughout. Differences between groups are described by Cohen's d effect size. The relationship between hind foot length and weight is explored by simple linear regression. All analyses are in R version 4.02 using RStudio version 1.3.1056


```{r, include = FALSE}
bonanza <- read_csv(here("data", "bonanza_hares.csv"))
```

### Preliminary Results
#### Juvenile hare trappings comparison and descriptive statistics

```{r, include = FALSE}
bonanza_year <- bonanza %>% # create new data set 
  mutate(date_new = mdy(date)) %>% 
  mutate(year = year(date_new)) %>% # with date converted to class "Date"
  filter(age == "j") %>% # only pull results where age = "j" (juvenile)
  select(year) %>% # select column `year`
  count(year) # counting total number of trappings per year

```

```{r}
bonanza_year_interactive <- ggplot(data = bonanza_year, aes(x = year, y = n)) +
  geom_line(color = "floralwhite") +
  labs (x = "Year",
        y = "Number of Trappings",
        title = "Juvenile Hares Trapped Per Year (1999 - 2012)") +
  theme(panel.border = element_blank(),
          panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.background = element_rect(fill = "darkslategrey"))

ggplotly(bonanza_year_interactive) # use `ggplotly()` to make the graph interactive!
```
**Figure 1. Total juvenile snowshoe hares trapped in Bonanza Creek LTER (1998 - 2012)** Trapping data was compiled from three sample sites of varying environment. Line tracks the total number of juvenile hares trapped each year.

```{r, include = FALSE}
stats_bonanza <- bonanza_year %>%
  group_by(year) %>%
  summarize(total_trappings = sum(n))

stats_bonanza %>%
  summarize(mean = mean(total_trappings),
            median = median(total_trappings),
            std_deviation = sd(total_trappings),
            maximum = max(total_trappings),
            minimum = min(total_trappings),
            sample_size = n())

```

Juvenile snowshoe hare weights (grams) were compared across three grid locations; sex was not included as a variable here, but will be investigated as a variable moving forward. Across all three sites, there is a visible decline in number of juvenile hares trapped between 1998 and 2012. Maximum number of hares trapped in a single year was `r max(stats_bonanza$total_trappings)` in 1999. Minimum number of hares trapped in a single year was `r min(stats_bonanza$total_trappings)` in 2010. Between 1999 and 2012 the mean and median number of juvenile hares trapped was `r mean(stats_bonanza$total_trappings)` $\pm$ `r sd(stats_bonanza$total_trappings)`, and `r median(stats_bonanza$total_trappings)`, respectively. To account for the fact that these counts are impacted by effort (e.g. number of days and traps in the study each year) you could standardize the population for future analyses by factoring in effort as a variable. Additionally, in other analyses that do not concern total number of trappings but focus on some other variables (e.g. weight and sex) you could choose to only compare a specific number of random samples. 

### Weight, sex, and grid location comparison

```{r}
juveniles <- bonanza %>% # create new data set 
  filter(age %in% "j") %>% # returning results for observations for juveniles
           select(weight, sex, age, grid) %>% # that selects for variables: sex, age, weight, and grid
  mutate(sex = case_when(sex == "m" ~ "Male", # for sex variable, rename m to male
                          sex == "f" ~ "Female"), # for sex variable, rename f to female
         grid = case_when(grid == "bonrip" ~ "Bonanza Riparian", # for grid location, rename bonrip to Bonanza Riparian
                          grid == "bonmat" ~ "Bonanza Mature", # for grid location, rename bonmat to Bonanza Mature
                          grid == "bonbs" ~ "Lowland Black Spruce")) # for grid location, rename bonbs to Lowland Black Spruce

ggplot(data = juveniles, aes(x = sex, y = weight)) + # plot juveniles data, comparing sex and weight
  geom_beeswarm(aes(color = sex)) + # make different sexes different colors
  scale_color_manual(values = c("cornflowerblue","chocolate"), name = "Sex") + # specifically female = blue, male = orange
  geom_boxplot(fill = NA, width = 0.2, outlier.color = NA) +
  stat_summary(fun=mean, 
               geom="point", 
               shape=20, 
               size=3, 
               color="black", 
               fill="black") +
  facet_wrap(~ grid) + # faceted by grid location
  theme_minimal() +
  labs(title = "Juvenile Hare Weight Distribution",
       x = "Sex",
       y = "Weight (g)")
```


**Figure 2. Juvenile snowshoe hare weight distribution Bonanza Creek LTER (1999 - 2012)** Weight observations (g) for juvenile snowshoe hares (male, and female) across three grid locations within Bonanza Creek LTER. Blue (female), orange (male), grey (NA) points indicate individual observations for weight (g) of juvenile snowshoe hares. Box endpoints indicate the 25^th^ and 75^th^ percentile values; the black point indicates the mean value for each sex and the black line indicates the median value for each sex. 

```{r}
ggplot(data = juveniles, aes(x = weight)) +
  geom_histogram(aes(fill = sex), # make histogram of hare weight distribution
                color = "black",
                 bins = 30,) +
  facet_grid(sex ~ grid) # faceted by sex and grid location
```

**Figure 3. Juvenile snowshoe hare weight distribution faceted by sex and grid location** Bars represent number of hares trapped. Colors represent hare sex red (female), blue (male), grey (NA).

Figure 1 and Figure 2 show the highest density of female and male juvenile snowshoe hares are found within the Bonanza Riparian site. The histograms in Figure 2 indicate that the data may not be normally distributed for any of the female or male populations sampled at any of the three sites. In particular, the distribution of male hares in the Bonanza Riparian site could be bimodal. 

### Juvenile weight comparison (male & female snowshoe hares)

```{r}

juveniles_stats <- juveniles %>%
  group_by(sex) %>%
  summarize(sample_size= n(),
            mean = round(mean(weight, na.rm = TRUE),2),
            std_deviation = round(sd(weight, na.rm = TRUE),2))

juveniles_stats %>%
  kable(col.names = c("Sex",
                      "Sample Size",
                      "Mean Size (g)",
                      "Standard Deviation"),
        caption = " **Table 1. Descriptive statistics (mean, standard deviation, and sample size) for male and female juvenile snowshoe hares (Bonanza Creek LTER).** Data was collected across three sites: Bonanza Riparian, Bonanza Mature, and Lowland Black Spruce. ") %>%
  kable_styling(full_width = FALSE)

```
```{r, include = FALSE}
ggplot(data = juveniles, aes(x = weight)) +
  geom_histogram(bins = 15) +
  facet_wrap(~ sex)

ggplot(data = juveniles, aes(sample = weight)) +
  geom_qq() +
  facet_wrap(~sex)
```
Investigative histograms & QQ plots for male and female juvenile snowshoe hares were relatively normal and symmetric. Histogram for juvenile males could be bimodal, histogram for NA values may be slightly skewed left. I feel OK about using a t-test for a means comparison.

```{r, include = FALSE}
juvenile_females <- juveniles %>% 
  filter(sex == "Female") %>% 
  drop_na(weight) %>%
  pull(weight)

juvenile_males <- juveniles %>%
  filter(sex == "Male") %>%
  drop_na(weight) %>%
  pull(weight)

juvenile_t <- t.test(juvenile_males, juvenile_females, na.rm = TRUE)

juvenile_cohen<- cohen.d(juvenile_males, juvenile_females, na.rm = TRUE)


male_mean <- mean(juvenile_males)
female_mean <- mean(juvenile_females)
male_sd <- sd(juvenile_males)
female_sd <- sd(juvenile_females)
```

On average, juvenile male snowshoe hares weigh more than juvenile female snowshoe hares (`r round(male_mean, 2)` $\pm$ `r round(male_sd, 2)` and `r round(female_mean, 2)` $\pm$ `r round(female_sd, 2)`, respectively; mean $\pm$ 1 standard deviation). While the absolute difference in means is `r round(male_mean, 2) - round(female_mean, 2)` g (a `r round((male_mean - female_mean)/((male_mean + female_mean)/2)*100, 2)`% difference), the difference in means is significant (Welch's two-sample t-test: t(`r round(juvenile_t$parameter, 2)`) = `r round(juvenile_t$statistic,2)`, p = `r juvenile_t$p.value`), and the effect size is small (Cohen's *d* = `r round(juvenile_cohen$estimate,2)`).

### Relationship between juvenile weight and hind foot length
#### Simple Linear Regression

```{r}
foot_weight <- bonanza %>%
  filter(age == "j") %>%
  select(age, hindft, weight)

ggplot(data = foot_weight, aes(x = hindft, y = weight)) +
  geom_point(color = "cornsilk4") +
  geom_smooth(method = lm, se = FALSE, color = "cornsilk2") +
  labs(title = "Hind Foot Length and Weight Comparison",
       x = "Hind Foot (mm)",
       y = "Weight (g)") +
  theme(panel.border = element_blank(),
          panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "black"))
```
```{r, include = FALSE}
foot_weight_lm <- lm(weight ~ hindft, data = foot_weight) # linear regression model

foot_weight_tidy <- tidy(foot_weight_lm) # tidy version of model

foot_weight_lm_glance <- glance(foot_weight_lm) # tidy version of model

plot(foot_weight_lm) # visualize linear model

foot_weight_lm_cor <- cor.test(foot_weight$weight, foot_weight$hindft) # calculate pearsons r correlation

foot_weight_lm_cor_tidy <- tidy(foot_weight_lm_cor)
```

**Figure 3.** Relationship between hind foot length (mm) and weight (g) for juvenile snowshoe hares. Linear model summary: $\beta$ ~1~ =  `r round(foot_weight_tidy$estimate[2], 2)` g mm^-1^, p < 0.001, R^2^ = `r round(foot_weight_lm_glance$r.squared, 2)`, Pearsons *r* = `r round(foot_weight_lm_cor_tidy$estimate, 2)`).

The relationship between hind foot length (mm) and weight (g) appears relatively linear across sexes (Figure 3). Simple linear regression revealed that hind foot length (mm) moderately and significantly predicts weight (p < 0.001, R^2^ = `r round(foot_weight_lm_glance$r.squared,2)`) with an average slope of $\beta$ = `r round(foot_weight_tidy$estimate[2], 2)` g mm^-1^ (i.e., for each one millimeter increase in hind foot length we expect an average increase in weight of `r round(foot_weight_tidy$estimate[2], 2)` g). Hind foot length and weight are moderately, and significantly, positively correlated (Pearson's *r* = `r round(foot_weight_lm_cor_tidy$estimate,2)`, p < 0.001). Diagnostic plots (not included) reveal somewhat normally distributed and mostly homoscedastic residuals. 

### Summary

Exploratory data analysis reveals the following initial findings: 

- Number of trappings of juvenile snowshoe hares from 1999 - 2012 vary and appear to steeply decline over time
- On average, male juvenile snowshoe hares weight more than females
- Hind foot length appears moderately linearly related with body mass across sexes; the relationship (by simple linear regression) is significant, with a slope (expected average increase in weight with a 1 mm increase in hind foot length) of `r round(foot_weight_tidy$estimate[2], 2)` g mm^-1^.

Next steps: 

- Compare hind foot lengths statistically (one-way ANOVA), including sex and/or trapping location (grid) as a covariate.
- Multiple linear regression to understand the relationship between trapping locaiton (grid), sex, hind foot length, and weight 

### Citations
Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative. https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14










